### 2014-11-18 poster/poster.tex
draft 1 of poster

%==Title, date and authors of the poster=======================================
\title
[CSSS 501 MLE Poster Session, 17 November 2014 -- http://github.com/pschmied/bikecounts-mle] % Conference
{ % Poster title
Weather and Seasonal Factors Affecting Daily Bicycle Counts at Fremont St. Bridge, Seattle
}

@@

\section{Model Specification}
Because some over dispersion was evident in the counts, we chose to
fit a negative binomial model. This model included the following variables

\begin{table}[htdp]
\begin{center}
\begin{tabular}{ll}
  \hline\hline
  Variable & Description \\
  \hline
  Count* & Number of bicycles / day \\
  Daylight hours & Length of daylight in hours \\
  Holiday & Was the day a holiday? \\
  UW & Was University of Washington in session? \\
  Max temp & Maximum temperature for the day \\
  Max precip & Daily max precip (inches) in any hour \\
  Day number & Sequentially numbered day of study \\
  Sat--Fri & Dummy variables for day of week \\
  \hline\hline
    & \multicolumn{1}{r}{*dependent variable}
\end{tabular}
\end{center}
\label{default}
\end{table}%

\section{Model Fit}

@@

@@

\subsection{Precipitation}

@@

\subsection{Day of the Week}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c3}
\caption{Simulated bicycle counts differences by day of week}
\label{fg:dow}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c4}
\caption{Effect of season (measured continuously as number of daylight hours)}
\label{default}
\end{center}
\end{figure}

\section{Conclusions}
%==============================================================================
%==End of content==============================================================
%==============================================================================

@@

\subsection{References}

\begin{thebibliography}{99}

\bibitem{ref1} J.~Doe, Article name, \textit{Phys. Rev. Lett.}

\bibitem{ref2} J.~Doe, J. Smith, Other article name, \textit{Phys. Rev. Lett.}

\bibitem{web} \url{http://www.google.pl}

\end{thebibliography}
%--End of references-----------------------------------------------------------

>>>>

%==Title, date and authors of the poster=======================================
\title
[CSSS 501 MLE Poster Session, 17 November 2014 -- http://github.com/pschmied/bikecounts-mle] % Conference
{ % Poster title
  Weather and Seasonal Factors Affecting Daily Bicycle Counts at
  Fremont St. Bridge, Seattle %
}

@@

\section{Model Specification}
Because over dispersion was evident in the counts, we chose to fit a
negative binomial model. This model includes the following variables,
and was fit using the \texttt{pscl::glm.nb()} function in R.

\begin{table}[htdp]
\begin{center}
\begin{tabular}{ll}
  \hline\hline
  Variable & Description \\
  \hline
  Count* & Number of bicycles / day \\
  Daylight hours & Length of daylight in hours \\
  Holiday & Was the day a holiday (recognized by UW)? \\
  UW & Was University of Washington in session? \\
  Max temp & Maximum temperature for the day \\
  Max precip & Daily max precip (inches) in any hour \\
  Day number & Sequentially numbered day of study \\
  Sat--Fri & Dummy variables for day of week \\
  \hline\hline
    & \multicolumn{1}{r}{*dependent variable}
\end{tabular}
\end{center}
\label{default}
\end{table}%

\section{Model Fit and Diagnostics}

As the actual-versus-predicted plot below shows, the model is a
reasonably good fit. However, we note several outliers falling far
from the $y=x$ line, perhaps indicating some other unaccounted source
of variation in the dataset.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.15\textwidth]{../fig/avpp1}
\caption{Actual versus predicted values from model}
\label{fg:avp}
\end{center}
\end{figure}

@@

As expected, the effect of temperature seems clear. Bicyclists at this
location appear to prefer warmer weather universally over cooler
weather. In a less temperate climate, we might expect bicycle counts
to fall again in very high temperature conditions.

@@

\subsection{Precipitation}
Similiarly, we see the strong negative effect of precipitation in our simulated
results.

@@

\subsection{Day of the Week}
As identified in our descriptive timeseries plot, these simulated
results confirm the presense of a weekly pattern in bicycling, with
much higher numbers of bicyclists during the week, falling off
somewhat toward Friday, and dropping precipitously on the weekends.
\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c3}
\caption{Simulated bicycle counts differences by day of week}
\label{fg:dow}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c4}
\caption{Effect of season (measured continuously as number of daylight
  hours)}
\label{fg:daylight}
\end{center}
\end{figure}

\subsection{General trend over time}
In our model we included a term that was a sequential numbering of the
days of the study period as to model a general trend over time. In the
visualization of the simulated results, we do see a general increase
in bicycle counts over this two year span.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.3\textwidth]{../fig/m1c5}
\caption{Simulated bicycle counts over study period}
\label{fg:trend}
\end{center}
\end{figure}

\section{Conclusions}
Over all, this model appears to be reasonably
predictive, with the results confirming our intuitions about factors
influencing bicycling rates. One major limitation, is that we still do
not account for the possibility that there is some timeseries
autoregression ocurring on a day-to-day basis. This is a problematic
limitation, in that we believe that the likelihood of bicycling one
day is very likely influenced by the choice to bicycle on the days
immediately preceeding.

%==============================================================================
%==End of content==============================================================
%==============================================================================

@@

% \subsection{References}

% \begin{thebibliography}{99}

% \bibitem{ref1} J.~Doe, Article name, \textit{Phys. Rev. Lett.}

% \bibitem{ref2} J.~Doe, J. Smith, Other article name, \textit{Phys. Rev. Lett.}

% \bibitem{web} \url{http://www.google.pl}

% \end{thebibliography}
%--End of references-----------------------------------------------------------
